\bf Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-15 appendix.qmd
Updated Figure 5 title

tsne_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
df_bin_pbmc <- algo_obj_pbmc$model_highd

@@

```{r}
#| echo: false
#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by RMSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters. The PBMC data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close. Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations. On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations. Therefore, @fig-pbmc-mse-umap b is more reasonable than @fig-pbmc-mse-umap a. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 90%

>>>>

tsne_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
#tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
df_bin_pbmc <- algo_obj_pbmc$model_highd

## Removed some bins manually to fit a better model
df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
  filter(!(hexID %in% c(559, 617, 677)))

tri_object <- tri_bin_centroids(
  centroids_data = df_bin_centroids_pbmc)

tr_from_to_df_pbmc <- gen_edges(tri_object = tri_object, a1 = algo_obj_pbmc$hb_obj$a1)
tr_from_to_df_pbmc <- tr_from_to_df_pbmc <- update_trimesh_index(trimesh_data = tr_from_to_df_pbmc)

@@

```{r}
#| echo: false
#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by RMSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters. The PBMC data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close. Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.04$, $b = 900/179 (30, 30)$) shows three-well separated clusters with big separations. On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.04$, $b = 1110/317 (30, 37)$) shows three-well separated clusters with small separations. Therefore, @fig-pbmc-mse-umap b is more reasonable than @fig-pbmc-mse-umap a. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 90%

### 2025-05-15 appendix.tex
Updated Figure 5 title

\caption{\label{fig-model-pbmc-author-proj}Compare the published \gD{}
layout (Figure~\ref{fig-pbmc-mse-umap} a) and the \gD{} layout selected
(Figure~\ref{fig-pbmc-mse-umap} b) by RMSE plot
(Figure~\ref{fig-pbmc-mse-umap}) from tSNE and UMAP with default
hyper-parameters. The PBMC data (\(n =  5858\)) has three clusters in
\(9\text{-}D\), where three clusters are close. Two \gD{} projection
from a tour on \(9\text{-}D\) of the model fit with
Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.06\),
\(b = 616/86 (22, 28)\)) shows three-well separated clusters with big
separations. On the other hand, the model fit with
Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.06\),
\(b = 704/227 (22, 32)\)) shows three-well separated clusters with small
separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
reasonable than Figure~\ref{fig-pbmc-mse-umap} a. Videos of the
langevitour animations are available at
\url{https://youtu.be/0cKX_HG_n0k} and
\url{https://youtu.be/KhJvsRtaX04} respectively.}

>>>>

\caption{\label{fig-model-pbmc-author-proj}Compare the published \gD{}
layout (Figure~\ref{fig-pbmc-mse-umap} a) and the \gD{} layout selected
(Figure~\ref{fig-pbmc-mse-umap} b) by RMSE plot
(Figure~\ref{fig-pbmc-mse-umap}) from tSNE and UMAP with default
hyper-parameters. The PBMC data (\(n =  5858\)) has three clusters in
\(9\text{-}D\), where three clusters are close. Two \gD{} projection
from a tour on \(9\text{-}D\) of the model fit with
Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.04\),
\(b = 900/179 (30, 30)\)) shows three-well separated clusters with big
separations. On the other hand, the model fit with
Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.04\),
\(b = 1110/317 (30, 37)\)) shows three-well separated clusters with
small separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
reasonable than Figure~\ref{fig-pbmc-mse-umap} a. Videos of the
langevitour animations are available at
\url{https://youtu.be/0cKX_HG_n0k} and
\url{https://youtu.be/KhJvsRtaX04} respectively.}
